---
title: "Titanic"
output: html_document
---
1.Read the data
```{r}
library(caret)

missing.types <- c('NA', '')
train.column.types <- c('integer',   # PassengerId
                        'factor',    # Survived 
                        'factor',    # Pclass
                        'character', # Name
                        'factor',    # Sex
                        'numeric',   # Age
                        'integer',   # SibSp
                        'integer',   # Parch
                        'character', # Ticket
                        'numeric',   # Fare
                        'character', # Cabin
                        'factor'     # Embarked
)

test.column.types <- train.column.types[-2]

train <- readData(Titanic.path, train.file.name, train.column.types, missing.types)
test <- readData(Titanic.path, test.file.name, test.column.types, missing.types)
test$Survived <- NA # add Survived column to test dataset and fill it out with NA.

combi <- rbind(train, test) # combine training and test set for further manipulation

```
2.Exploratory Data abalysis
```{r}
str(train)
str(test)
summary(train)
summary(test)
```
Visualize the relationship between features and response

Check the relation between Sex and Survived
```{r}
library(ggplot2)
p <- ggplot(train, aes(x=Survived,fill = Sex) )+geom_bar()
p
```
Consistent with our prior, the female is more likely to survive.


Check the relation between Age and Survive.
```{r}
#histogram
p  <- ggplot(train, aes(x=Age, fill = Survived)) + geom_histogram()
p
#find missing value in Age
x <- which(is.na(train$Age))
length(x)
p  <- ggplot(train[!is.na(train$Age), ], aes(x=Age, fill = Survived)) + geom_histogram()
p
#density plot
#-which(is.na) = !is.na
p1 <- ggplot(train[-which(is.na(train$Age)), ], aes(x=Age, fill=Survived)) + geom_density(alpha=0.6)+ facet_grid(.~Sex);
p1
p1 <- ggplot(train[!is.na(train$Age), ], aes(x=Age, fill=Survived)) + geom_density(alpha=0.6)+ facet_grid(.~Sex);
p1
```
Check the relation between Fare and Survive.
```{r}
p <- ggplot(train, aes(x=Fare, fill=Survived)) + geom_histogram()+facet_grid(~Pclass)
p
#Apperently,peolple who survived afforded highe fare for the ticked, which indicates the higher fare may correspond to the cabin with better escaping route.
```

```{r}
#prop.table can create proportion, 
#margin = 2, divide by the sum of col
#margin = 1, divide by the sum of row
t <- table(train$Survived,train$Pclass)
t
prop.table(t, margin = 2)

```
We can see that Proportion of Surviers in Pclass 1 are higher than Pclass2 and Pclass3.


Check the relation between Embarked and Survived.
```{r}
p <- ggplot(train[!is.na(train$Embarked),] , aes(x= Embarked , fill = Survived)) + geom_bar()
p
```
```{r}
p <- ggplot(train , aes(x= Embarked , fill = Survived)) + geom_bar()
p

```
3. Data Clearing and Feature Engineering
```{r}
title.extract <- function(x){
    strsplit(x, split = "[,.]")[[1]][2]
}

combi$Title <- sapply(combi$Name, FUN = title.extract)
combi$Title <- sub(" ","",x = combi$Title)

combi$Title <- as.factor(combi$Title)
combi$Surname <- sapply(combi$Name, FUN = function(x){
    strsplit(x, split = "[,.]")[[1]][1]
})

```

Use SibSp and Parch. Create a new feature FamilySize which equals SibSq + Parch +1
```{r}
combi$FamilySize <- combi$SibSp + combi$Parch + 1
combi$FamilyID <- paste(as.character(combi$FamilySize), combi$Surname, sep = "")
combi$FamilyID[combi$FamilySize <= 2] <- "Small"

famIDs <- data.frame(table(combi$FamilyID))

famIDs <- famIDs[famIDs$Freq <= 2,] 

combi$FamilyID[combi$FamilyID %in% famIDs$Var1] <- "Small"
combi$FamilyID <- as.factor(combi$FamilyID)

```

